Alleviating Topology Mismatch in Distributed Overlay Networks: A Survey

### 2014-01-27 intro.tex
2012.12.23 (Mema) Text reworked.

Two ``killer'' applications of the peer-to-peer paradigm are, but not limited
to, file-sharing and telephony. P2P file-sharing was originally introduced by
Napster (see Section~\ref{section:background}), which by 2001 was widely
acknowledged as the ``fastest growing Internet application ever" with 26 million
users sharing 80 million songs at the height of its popularity. Other similar
applications quickly followed suit including Gnutella and its descendant,
Limewire, enjoying 3 million concurrently connected peers by the end of 2006
and BitTorrent, connecting over 150 million users. P2P telephony has seen
explosive growth with the advent of Skype. Since its introduction in 2003,
Skype has become extremely popular with more than 650 million users in 2011
\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
March 2012 \cite{skypesymusers}.

@@

One of the fundamental issues that defines the efficiency of an overlay network
is how well the overlay maps to the underlying network topology on which it
rests. Consider two nodes\footnote{In the case of peer-to-peer networks,
typically the participating nodes are user PCs sitting at the edge of the
Internet.} that share a link in an overlay network which maps to multiple hops
in the underlying (say IP) topology. If the traffic patterns of the application
are such that traffic flows heavily over the overlay link between the two nodes,
then it is beneficial to construct the overlay topology such that the number of
underlying IP links between these two nodes is minimized.

If the overlay network is constructed such that it does not match the underlying
topology well, the resulting \emph{topology mismatch} (See
Section~\ref{section:background} for a more formal definition) can result in two
significant problems. First, the performance of the application per se, can be
adversely affected since application traffic must flow over a larger,
redundunt, number of physical hops resulting in poor user experience like
extensive search latency or jitter. Second, can cumulatively degrade the
performance of the whole underlying infrastructure, effecting other applications
as well. For example, studies have shown that the extreme popularity of P2P
applications amply deployed everywhere nowadays, result in them contributing the
largest portion of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some reporting that more than 60\% of it, is
P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others claiming that it
is expected to reach almost 8 petabytes per month by the end of
2012~\cite{multinteligence}. This is a significant burden for Internet Service
Providers (ISPs) who must handle all this traffic to destination nodes sitting
at the edge of the Internet. If the overlay topology of a peer-to-peer system is
not efficiently constructed, then this burden to the Internet's backbone
infrastructure could be even greater. For example, traffic might have to flow
back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node. Thus, it is fundamentally
important that P2P overlay networks are constructed and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several researchers have published profusely on the topology
mismatch problem.  In this paper, we survey this research area and provide a
taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. Ultimately, our goal
is to help readers sift through the vast amount of literature, to help them
understand the advantages and disadvantages of each of each work, and to provide
them with enough perspective so that when the need arises, they are able to
select amongst the different approaches for their particular application.

The rest of the paper is organized as follows: In
Section~\ref{section:background}, we provide background information on
peer-to-peer overlay architectures including centralized, decentralized
unstructured and decentralized structured peer-to-peer systems. We also formally
define the problem of topology mismatch between an overlay and its underlying
physical network and elaborate on the rationale behind our coarse grained
evaluation of the techniques presented throughout this survey. In
Sections~\ref{section:unstructured} and \ref{section:structured}, we present the
various research efforts for decentralized unstructured and decentralized
structured peer-to-peer systems, respectively. Finally, with
Section~\ref{section:conclusion}, we conclude the paper.

>>>>

Two ``killer'' applications of the peer-to-peer paradigm are 
file-sharing and telephony. P2P file-sharing was originally introduced by
Napster (see Section~\ref{section:background}), which by 2001 was widely
acknowledged as the ``fastest growing Internet application ever" with 26 million
users sharing 80 million songs at the height of its popularity. Other file-sharing 
applications quickly followed suit including Gnutella and its descendant,
Limewire, enjoying 3 million concurrently connected peers by the end of 2006
and BitTorrent, connecting over 150 million users. P2P telephony, on the
other hand, has seen
explosive growth with the advent of Skype. Since its introduction in 2003,
Skype has become extremely popular with more than 650 million users in 2011
\cite{skypetotalusers} and 35 million skypers concurrently online on the 5th of
March 2012 \cite{skypesymusers}.

@@

One of the fundamental issues that defines the efficiency of an overlay network
is how well the overlay maps to the underlying network topology on which it
rests. Consider two nodes\footnote{In the case of peer-to-peer networks,
typically the participating nodes are user PCs sitting at the edge of the
Internet.} that share a link in an overlay network which maps to multiple hops
in the underlying IP topology. If the traffic patterns of the application
are such that traffic flows heavily over the overlay link between the two nodes,
then it is beneficial to construct the overlay topology such that the number of
underlying IP links between these two nodes is minimized.

If the overlay network is constructed such that it does not match the underlying
topology well, the resulting \emph{topology mismatch} 
can result in two
significant problems. First, the performance of the application per se, can be
adversely affected since application traffic must flow over a larger,
redundunt, number of physical hops resulting in poor user experience like
extensive search latency or jitter. Second, other applications running
on the underlying network infrastructure are also adversely affected.
Studies have shown that highly popular P2P applications contribute the largest
portion of overall Internet traffic
\cite{seroiu_analysiscds_2002,sen_analyzep2ptraffic_2004,krp_ispfear_2005}, with
some studies reporting that more than 60\% of Internet traffic is
P2P-related~\cite{cachelogic,ipoque2007,ipoque2009} and others predicting that it
will reach almost
8 petabytes per month by the end of
2012~\cite{multinteligence}. This is a significant burden for Internet Service
Providers (ISPs) who must route all of this traffic to destination nodes sitting
at the edge of the Internet. If the overlay topology of a peer-to-peer system is
not efficiently constructed, then the burden on the Internet's backbone
infrastructure could be even greater because traffic might have to flow
back and forth several times between two neighboring ISPs as it travels from its
source node in the overlay to its destination node. Thus, it is fundamentally
important that P2P overlay networks are constructed and maintained so that
their topology matches as closely as possible the underlying IP topology.

For over a decade, several researchers have published profusely on the topology
mismatch problem.  In this paper, we survey this research area and provide a
taxonomy of the proposed solutions. We point out synergies as well as
similarities and differences in the published approaches. Ultimately, our goal
is to help readers sift through the vast amount of literature, to help them
understand the advantages and disadvantages of each work, and to provide
them with enough perspective so that when the need arises, they are able to
select amongst the different approaches for their particular application.

The rest of the paper is organized as follows: In
Section~\ref{section:background}, we provide background information on
peer-to-peer overlay architectures including centralized, decentralized
unstructured and decentralized structured peer-to-peer systems. We also formally
define the problem of topology mismatch between an overlay and its underlying
physical network and present the rationale behind our 
evaluation of the techniques presented throughout this survey. In
Sections~\ref{section:unstructured} and \ref{section:structured}, we present the
various research efforts for decentralized unstructured and decentralized
structured peer-to-peer systems, respectively. Finally, in 
Section~\ref{section:conclusion}, we conclude the paper.

### 2014-01-27 mard-survey.tex
2012.12.23 (Mema) Text reworked.

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The
overlay network abstraction on top of a best-effort infrastructure provides a
powerful tool to realize a plethora of ever-wanted exotic features like
anonymity, high availability and robustness, load balancing, quality-of-service
(QoS) at the application layer and many more. Unfortunately weaknesses of early
effectuations constrained applications from unleashing the full potential of the
paradigm. One of them is the topology mismatch problem between the overlay and
the physical underlying network, a problem that imposes a huge amount of
unnecessary stress to the network resources and, thus, impede applications from
scaling. In this survey we investigate on the recent research developments
towards the alleviation of this problem, in both unstructured and structured
overlay architectures. We visit the various algorithms available in the
literature, focusing on the presentation of how each of them tries to tackle the
problem. Additionally, we provide a finer grained categorization of the
approaches based on the generic type of methodology they use in order to
accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the
algorithms and offer pictorial comparisons for them in order to further
strengthen and organize understanding of the information and knowledge cited, in
an attempt to make this survey a one-stop-shop reference for anyone interested
in the topic, from researchers and students, to system architects and
developers.
\end{abstract}

>>>>

\begin{abstract}
Peer-to-peer (P2P) systems have enjoyed immense attention and have been
widely deployed on the Internet for well over a decade.  P2P systems 
are implemented via
an overlay network abstraction on top of the best-effort IP infrastructure
of the Internet. These systems offer a plethora
of desirable features to distributed applications that are built on top of them
including anonymity, high availability and robustness, load balancing,
quality of service, scalability, and others.  Unfortunately, weaknesses of
early deployments of P2P systems prevented applications from leveraging 
these features.  One major weakness, identified early on, is the topology
mismatch problem between the overlay network and the underlying IP topology.
This mismatch can impose an extraordinary amount of unnecessary stress on
network resources and can adversely affect the scalability of applications
built on top of P2P systems.  
In this paper, we survey over a decade's worth of research efforts
aimed at alleviating the topology mismatch problem, in both unstructured
and structured P2P systems.   
We provide a fine-grained categorization of the
proposed solutions and discuss the advantages, disadvantages, and novelties
introduced by each solution.  Finally, we provide an analysis and pictorial 
comparisons 
of the proposed solutions in an attempt to provide a quick one-stop-shop
reference for anyone interested in the topic, from researchers and students,
to system architects and developers.

\end{abstract}
